ABM simulating short-term mating behaviors

### 2022-06-01 submission/manuscript.Rmd
added texts to explain in-poll differences and added hypos

      formatted <- ifelse(
        abs(x - round(x)) < .Machine$double.eps,
        # treat as integer
        format(x, digits=0, big.mark = ","),
        # treat as floating point number
        format(x, digits=1, nsmall=2, big.mark = ",")
        )

@@

As a starting point, we took gender differences in mating preferences as conceptualized by the sexual strategy theory [@buss_sexual_1993], which is repeatedly supported by empirical investigations [e.g., @schmitt_universal_2003; @walter_sex_2020]. Using an agent-based model, we investigated whether men and women's differential preferences for short-term mating would result in gender differences in short-term mating behaviors, and if so, then under what conditions they would result in differences in the number of short-term mating experiences and short-term mates.

@@

Gender differences in mating preferences have been studied in the light of sexual strategy theory [@buss_sexual_1993]. It posits that since the minimum obligatory investment that men must devote to their offspring (contribution of sperm through one sexual act) is lower than that of women (gestation, labor, and lactation), men tend to be relatively more interested in short-term mating compared to women. This is because the number of offspring women can produce is limited and cannot be increased by mating with a large number of men, whereas the reverse is the case for men. As a result, men are predicted to not only have a greater interest in short-term mating, but also desire a larger number of short-term mates in a given period and be less selective with respect to accepting a potential mate [@buss_sexual_1993].

@@

## Constraints on men's preferences for short-term mating
Since heterosexual sex in most cases involves one man and one woman, men's relatively high interest in short-term mating can be constrained by women's preference for long-term relationships [@archer_reality_2019; @symons_evolution_1979]. This is because men’s short-term mating preferences – with the exception of sexual violence – can only translate into behaviors when there are women willing to have sex with them. When a new short-term mating encounter occurs, it counts towards both men’s total number of short-term mating encounters as well as women’s, and both men and women have new short-term mates. Therefore, the total number of short-term mating experiences and short-term mates must be equal between heterosexual men and women at the population level. However, it is important to note that although heterosexual men’s total number of short-term mating is equal to that of women, this does not necessarily mean that the total number of men and women who have ever had short-term mating must be equal. It is possible that there is a smaller proportion of men (vs. the proportion of women) who contribute to the total number of short-term mating. Following this line of reasoning, we would expect to observe no gender difference in short-term mating behaviors among heterosexual individuals even when there are gender differences in mating preferences.

@@

We also modeled forming long-term relationships as a background process. A pair of individuals had a possibility to commit to a long-term relationship upon meeting each other. We modeled forming long-term relationships as a behavioral outcome and ignored the psychological processes behind it, although we recognize gender differences may also exist in long-term mating strategy [@buss_sexual_1993].

@@

We formulated the following hypotheses in the present study: assuming gender differences in preferences for short-term mating, 1) there would be gender differences in short-term mating behaviors among gay men and lesbian women, with gay men being engaged in more such behaviors, and 2) gay men would engage in more short-term mating behaviors as compared to heterosexual men.

@@

## Gender differences in short-term mating behaviors

@@

The results from Experiment 1 showed that when there were gender differences in preferences for short-term mating, there was no gender difference in the average number of short-term mating experiences among heterosexual individuals (*M~men~* = `r m_exp_all.mean_011`, *SD~men~* = `r m_exp_all.sd_011`; *M~women~* = `r f_exp_all.mean_011`, *SD~women~* = `r f_exp_all.sd_011`), Cohen's *d* = `r exp_all.d_011`. Nor was there a difference in the average number of short-term mates (*M~men~* = `r m_partner_all.mean_011`, *SD~men~* = `r m_partner_all.sd_011`; *M~w~* = `r f_partner_all.mean_011`, *SD~women~* = `r f_partner_all.sd_011`), Cohen's *d* = `r partner_all.d_011` (Figure 1).

@@

```{r sec1.3}
# the number of experiences, difference*2, heterosexual in the mating pool
m_exp_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_ip.mean)
m_exp_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_ip.sd)
f_exp_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_ip.mean)
f_exp_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_ip.sd)
exp_ip.parameter_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.parameter)
exp_ip.statistic_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.statistic)
exp_ip.p.value_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.p.value)
exp_ip.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.d) %>% abs()

>>>>

      formatted <- ifelse(
        abs(x - round(x)) < .Machine$double.eps & x != 0,
        # treat as integer
        format(x, digits=0, big.mark = ","),
        # treat as floating point number
        format(x, digits=1, nsmall=2, big.mark = ",")
        )

@@

As a starting point, we took gender differences in mating preferences as conceptualized by the sexual strategy theory [@buss_sexual_1993; @buss2019mate], which is repeatedly supported by empirical investigations [e.g., @schmitt_universal_2003; @walter_sex_2020]. Using an agent-based model, we investigated whether men and women's differential preferences for short-term mating would result in gender differences in short-term mating behaviors, and if so, then under what conditions they would result in differences in the number of short-term mating experiences and short-term mates.

@@

Gender differences in mating preferences have been studied in the light of sexual strategy theory [@buss_sexual_1993; @buss2019mate]. It posits that since the minimum obligatory investment that men must devote to their offspring (contribution of sperm through one sexual act) is lower than that of women (gestation, labor, and lactation), men tend to be relatively more interested in short-term mating compared to women. This is because the number of offspring women can produce is limited and cannot be increased by mating with a large number of men, whereas the reverse is the case for men. As a result, men are predicted to not only have a greater interest in short-term mating, but also desire a larger number of short-term mates in a given period and be less selective with respect to accepting a potential mate [@buss_sexual_1993].

@@

## Constraints on men's preferences for short-term mating

Since heterosexual sex in most cases involves one man and one woman, men's relatively high interest in short-term mating can be constrained by women's preference for long-term relationships [@archer_reality_2019; @symons_evolution_1979]. This is because men’s short-term mating preferences – with the exception of sexual violence – can only translate into behaviors when there are women willing to have sex with them. When a new short-term mating encounter occurs, it counts towards both men’s total number of short-term mating encounters as well as women’s, and both men and women have new short-term mates. Therefore, the total number of short-term mating experiences and short-term mates must be equal between heterosexual men and women at the population level. In a closed population with equal sex ratio, it follows that the average number of experiences and mates should also be equal between men and women. Following this line of reasoning, we would expect to observe no gender difference in short-term mating behaviors among heterosexual individuals even when there were gender differences in mating preferences.

However, it is important to note that although heterosexual men’s total number of short-term mating is equal to that of women, this does not necessarily mean that the total number of men and women who have ever had short-term mating must be equal. Since women have higher standards for short-term mates than men [@buss_sexual_1993], it is possible that the proportion of individuals who meet a potential partner's standards is lower among men than among women, assuming both men and women's desirability values follow the same distribution. Consequently, there may be a smaller proportion of men (vs. the proportion of women) who participate in the mating pool and contribute to the total number of short-term mating. For example, suppose that there is a population of 20 individuals with an equal sex ratio. There are 8 men and 8 women with a desirability value greater than x~1~ and 3 men and 3 women with a value greater than x~2~ (x~1~ < x~2~). Suppose that x~1~ and x~2~ happen to be men and women's standards for short-term mates, then in a long enough period of time, the 8 women who meet men's standards and the 3 men who meet women's standards will have had short-term mating. Since the total number of short-term mating behaviors is the same between men and women, we would expect that in a population of individuals who had any short-term mating, the average number of short-term mating experiences and mates was higher among men than women (e.g., in the aforementioned example, the number is n/3 for men and n/8 for women, where n can be the total number of short-term mating experiences or the number of short-term mates).

@@

We also modeled forming long-term relationships as a background process. A pair of individuals had a possibility to commit to a long-term relationship upon meeting each other. We modeled forming long-term relationships as a behavioral outcome and ignored the psychological processes behind it, although we recognize gender differences may also exist in long-term mating strategy [@buss_sexual_1993; @buss2019mate].

@@

We formulated the following hypotheses in the present study: assuming gender differences in preferences for short-term mating, 1) there would be no gender differences in short-term mating behaviors among all heterosexual men and women, 2) a smaller proportion of heterosexual men had ever engaged into short-term mating compared to the proportion of heterosexual women, 3) in a population of heterosexual individuals who had any short-term mating, men had higher average number of short-term mating behaviors than women, 4) there would be gender differences in short-term mating behaviors among gay men and lesbian women, with gay men being engaged in more such behaviors, and 5) gay men would engage in more short-term mating behaviors compared to heterosexual men.

@@

## Gender differences in short-term mating behaviors in the full population

@@

The results from Experiment 1 showed that when there were gender differences in preferences for short-term mating, there was no gender difference in the average number of short-term mating experiences among heterosexual individuals (*M~men~* = `r m_exp_all.mean_011`, *SD~men~* = `r m_exp_all.sd_011`; *M~women~* = `r f_exp_all.mean_011`, *SD~women~* = `r f_exp_all.sd_011`), Cohen's *d* = `r exp_all.d_011`. Nor was there a difference in the average number of short-term mates (*M~men~* = `r m_partner_all.mean_011`, *SD~men~* = `r m_partner_all.sd_011`; *M~women~* = `r f_partner_all.mean_011`, *SD~women~* = `r f_partner_all.sd_011`), Cohen's *d* = `r partner_all.d_011` (Figure 1).

@@

## Heterosexual individuals in the mating pool

```{r sec2.2}
# the number of experiences, difference*2, heterosexual in the mating pool
m_exp_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_ip.mean)
m_exp_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(m_exp_ip.sd)
f_exp_ip.mean_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_ip.mean)
f_exp_ip.sd_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(f_exp_ip.sd)
exp_ip.parameter_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.parameter)
exp_ip.statistic_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.statistic)
exp_ip.p.value_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.p.value)
exp_ip.d_011 <-
  results %>% filter(hls == "0_1_1") %>% pull(exp_ip.d) %>% abs()

### 2022-06-01 submission/manuscript.tex
added texts to explain in-poll differences and added hypos

As a starting point, we took gender differences in mating preferences as
conceptualized by the sexual strategy theory (Buss \& Schmitt, 1993),
which is repeatedly supported by empirical investigations (e.g.,
Schmitt, 2003; Walter et al., 2020). Using an agent-based model, we
investigated whether men and women's differential preferences for
short-term mating would result in gender differences in short-term
mating behaviors, and if so, then under what conditions they would
result in differences in the number of short-term mating experiences and
short-term mates.

@@

Gender differences in mating preferences have been studied in the light
of sexual strategy theory (Buss \& Schmitt, 1993). It posits that since
the minimum obligatory investment that men must devote to their
offspring (contribution of sperm through one sexual act) is lower than
that of women (gestation, labor, and lactation), men tend to be
relatively more interested in short-term mating compared to women. This
is because the number of offspring women can produce is limited and
cannot be increased by mating with a large number of men, whereas the
reverse is the case for men. As a result, men are predicted to not only
have a greater interest in short-term mating, but also desire a larger
number of short-term mates in a given period and be less selective with
respect to accepting a potential mate (Buss \& Schmitt, 1993).

@@

Since heterosexual sex in most cases involves one man and one woman,
men's relatively high interest in short-term mating can be constrained
by women's preference for long-term relationships (Archer, 2019; Symons,
1979). This is because men's short-term mating preferences -- with the
exception of sexual violence -- can only translate into behaviors when
there are women willing to have sex with them. When a new short-term
mating encounter occurs, it counts towards both men's total number of
short-term mating encounters as well as women's, and both men and women
have new short-term mates. Therefore, the total number of short-term
mating experiences and short-term mates must be equal between
heterosexual men and women at the population level. However, it is
important to note that although heterosexual men's total number of
short-term mating is equal to that of women, this does not necessarily
mean that the total number of men and women who have ever had short-term
mating must be equal. It is possible that there is a smaller proportion
of men (vs.~the proportion of women) who contribute to the total number
of short-term mating. Following this line of reasoning, we would expect
to observe no gender difference in short-term mating behaviors among
heterosexual individuals even when there are gender differences in
mating preferences.

@@

We also modeled forming long-term relationships as a background process.
A pair of individuals had a possibility to commit to a long-term
relationship upon meeting each other. We modeled forming long-term
relationships as a behavioral outcome and ignored the psychological
processes behind it, although we recognize gender differences may also
exist in long-term mating strategy (Buss \& Schmitt, 1993).

@@

\hypertarget{gender-differences-in-short-term-mating-behaviors}{%
\subsection{Gender differences in short-term mating
behaviors}\label{gender-differences-in-short-term-mating-behaviors}}

The results from Experiment 1 showed that when there were gender
differences in preferences for short-term mating, there was no gender
difference in the average number of short-term mating experiences among
heterosexual individuals (\emph{M\textsubscript{men}} = 1.43,
\emph{SD\textsubscript{men}} = 0.24; \emph{M\textsubscript{women}} =
1.43, \emph{SD\textsubscript{women}} = 0.24), Cohen's \emph{d} = 0. Nor
was there a difference in the average number of short-term mates
(\emph{M\textsubscript{men}} = 0.80, \emph{SD\textsubscript{men}} =
0.11; \emph{M\textsubscript{w}} = 0.80, \emph{SD\textsubscript{women}} =
0.11), Cohen's \emph{d} = 0 (Figure 1).

@@

@@

>>>>

As a starting point, we took gender differences in mating preferences as
conceptualized by the sexual strategy theory (Buss \& Schmitt, 1993;
Buss \& Schmitt, 2019), which is repeatedly supported by empirical
investigations (e.g., Schmitt, 2003; Walter et al., 2020). Using an
agent-based model, we investigated whether men and women's differential
preferences for short-term mating would result in gender differences in
short-term mating behaviors, and if so, then under what conditions they
would result in differences in the number of short-term mating
experiences and short-term mates.

@@

Gender differences in mating preferences have been studied in the light
of sexual strategy theory (Buss \& Schmitt, 1993; Buss \& Schmitt,
2019). It posits that since the minimum obligatory investment that men
must devote to their offspring (contribution of sperm through one sexual
act) is lower than that of women (gestation, labor, and lactation), men
tend to be relatively more interested in short-term mating compared to
women. This is because the number of offspring women can produce is
limited and cannot be increased by mating with a large number of men,
whereas the reverse is the case for men. As a result, men are predicted
to not only have a greater interest in short-term mating, but also
desire a larger number of short-term mates in a given period and be less
selective with respect to accepting a potential mate (Buss \& Schmitt,
1993).

@@

Since heterosexual sex in most cases involves one man and one woman,
men's relatively high interest in short-term mating can be constrained
by women's preference for long-term relationships (Archer, 2019; Symons,
1979). This is because men's short-term mating preferences -- with the
exception of sexual violence -- can only translate into behaviors when
there are women willing to have sex with them. When a new short-term
mating encounter occurs, it counts towards both men's total number of
short-term mating encounters as well as women's, and both men and women
have new short-term mates. Therefore, the total number of short-term
mating experiences and short-term mates must be equal between
heterosexual men and women at the population level. In a closed
population with equal sex ratio, it follows that the average number of
experiences and mates should also be equal between men and women.
Following this line of reasoning, we would expect to observe no gender
difference in short-term mating behaviors among heterosexual individuals
even when there were gender differences in mating preferences.

However, it is important to note that although heterosexual men's total
number of short-term mating is equal to that of women, this does not
necessarily mean that the total number of men and women who have ever
had short-term mating must be equal. Since women have higher standards
for short-term mates than men (Buss \& Schmitt, 1993), it is possible
that the proportion of individuals who meet a potential partner's
standards is lower among men than among women, assuming both men and
women's desirability values follow the same distribution. Consequently,
there may be a smaller proportion of men (vs.~the proportion of women)
who participate in the mating pool and contribute to the total number of
short-term mating. For example, suppose that there is a population of 20
individuals with an equal sex ratio. There are 8 men and 8 women with a
desirability value greater than x\textsubscript{1} and 3 men and 3 women
with a value greater than x\textsubscript{2} (x\textsubscript{1}
\textless{} x\textsubscript{2}). Suppose that x\textsubscript{1} and
x\textsubscript{2} happen to be men and women's standards for short-term
mates, then in a long enough period of time, the 8 women who meet men's
standards and the 3 men who meet women's standards will have had
short-term mating. Since the total number of short-term mating behaviors
is the same between men and women, we would expect that in a population
of individuals who had any short-term mating, the average number of
short-term mating experiences and mates was higher among men than women
(e.g., in the aforementioned example, the number is n/3 for men and n/8
for women, where n can be the total number of short-term mating
experiences or the number of short-term mates).

@@

We also modeled forming long-term relationships as a background process.
A pair of individuals had a possibility to commit to a long-term
relationship upon meeting each other. We modeled forming long-term
relationships as a behavioral outcome and ignored the psychological
processes behind it, although we recognize gender differences may also
exist in long-term mating strategy (Buss \& Schmitt, 1993; Buss \&
Schmitt, 2019).

@@

\hypertarget{gender-differences-in-short-term-mating-behaviors-in-the-full-population}{%
\subsection{Gender differences in short-term mating behaviors in the
full
population}\label{gender-differences-in-short-term-mating-behaviors-in-the-full-population}}

The results from Experiment 1 showed that when there were gender
differences in preferences for short-term mating, there was no gender
difference in the average number of short-term mating experiences among
heterosexual individuals (\emph{M\textsubscript{men}} = 1.43,
\emph{SD\textsubscript{men}} = 0.24; \emph{M\textsubscript{women}} =
1.43, \emph{SD\textsubscript{women}} = 0.24), Cohen's \emph{d} = 0.00.
Nor was there a difference in the average number of short-term mates
(\emph{M\textsubscript{men}} = 0.80, \emph{SD\textsubscript{men}} =
0.11; \emph{M\textsubscript{w}} = 0.80, \emph{SD\textsubscript{women}} =
0.11), Cohen's \emph{d} = 0.00 (Figure 1).

@@

\hypertarget{heterosexual-individuals-in-the-mating-pool}{%
\subsection{Heterosexual individuals in the mating
pool}\label{heterosexual-individuals-in-the-mating-pool}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-buss2019mate}{}}%
Buss, D. M., \& Schmitt, D. P. (2019). Mate preferences and their
behavioral manifestations. \emph{Annual Review of Psychology},
\emph{70}, 77--110.
